---
title: "R Notebook"
output: html_notebook
---

# Step 0 - Operationalize your analysis 
This is where you make explicit what your measuring (accuracy ≠ precision) and what.

The validity of the models is informed by theory, exisiting literature, intuition and observations.

# Working with means
Imagine that you're working with F0 measurements made with a population of speakers. 

```{r}
data <- read_csv("data/F0.csv")
```
# Is this valid data?

```{r}

data |> 
    ggplot(aes(x = 1, y = F0)) +
    geom_jitter()

```


```{r}

data |> 
    ggplot(aes(x = 1, y = F0, color = gender)) +
    geom_jitter()
    

```

```{r}

data |> 
    ggplot(aes(x = F0)) +
    geom_density()

```

```{r}

pluc

data |> 
    ggplot(aes(x = 1, y = F0, color = gender)) +
    geom_jitter() +
    geom_hline(color = "blue", yintercept = data |> filter(gender == "M") |> pull(F0) |> mean()) +
    geom_hline(color = "red", yintercept = data |> filter(gender == "F") |> pull(F0) |> mean())
```
## Quantitative analysis 

We have a good idea that the two populations have different F0s. How do we go about "proving" that is the case?

A t test can be very useful when you are compating measurements in a populatio

n. 
```{r}
vector_m <- data |> filter(gender == "M") |> pull(F0)
vector_f <- data |> filter(gender == "F") |> pull(F0)


t.test(vector_m, vector_f)

```

# Classification problem

Here's the problem. I asked a bunch of people to judge whether a sentence they hear was meant to be interpreted as literal or idiomatic. How do they do?

Here's why operationalizing is very important. 

```{r}

data <- read_csv("data/perception.csv")

glimpse(data)

data |> 
    ggplot() +
    geom_col(aes(x = n_success, y = expected), fill = "red", alpha = 0.25) +
    geom_col(aes(x = n_success, y = observed), fill = "blue", alpha = 0.25)

```

